Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations

### 2016-12-06 paper.tex
Huge commit

\documentclass[11pt, oneside]{article}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{ptex2tex, minted}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

\definecolor{gray}{gray}{0.97}
\colorlet{commentcolour}{green!50!black}
\colorlet{stringcolour}{red!60!black}
\colorlet{keywordcolour}{magenta!90!black}
\colorlet{exceptioncolour}{yellow!50!red}
\colorlet{commandcolour}{blue!60!black}
\colorlet{numpycolour}{blue!60!green}
\colorlet{literatecolour}{magenta!90!black}
\colorlet{promptcolour}{green!50!black}
\colorlet{specmethodcolour}{violet}
\colorlet{indendifiercolour}{green!70!white}

@@

\lstset{
numbers=none,
aboveskip=1ex,
belowskip=1ex,
basicstyle=\ttfamily\footnotesize,
}

\lstdefinestyle{pythonstyle}{
%%\lstset{
%%keepspaces=true,
language=python,
showtabs=true,
tab=,
tabsize=2,
basicstyle=\ttfamily\footnotesize,%\setstretch{.5},
stringstyle=\color{stringcolour},
showstringspaces=false,
alsoletter={1234567890},
otherkeywords={\ , \}, \{, \%, \&, \|},
keywordstyle=\color{keywordcolour}\bfseries,
emph={and,break,class,continue,def,yield,del,elif ,else,%
except,exec,finally,for,from,global,if,import,in,%
lambda,not,or,pass,print,raise,return,try,while,assert},
emphstyle=\color{blue}\bfseries,
emph={[2]True, False, None},
emphstyle=[2]\color{keywordcolour},
emph={[3]object,type,isinstance,copy,deepcopy,zip,enumerate,reversed,list,len,dict,tuple,xrange,append,execfile,real,imag,reduce,str,repr},
emphstyle=[3]\color{commandcolour},
emph={Exception,NameError,IndexError,SyntaxError,TypeError,ValueError,OverflowError,ZeroDivisionError},
emphstyle=\color{exceptioncolour}\bfseries,
%upquote=true,
morestring=[s]{"""}{"""},
morestring=[s]{'''}{'''},
commentstyle=\color{commentcolour}\slshape,
emph={[4]1, 2, 3, 4, 5, 6, 7, 8, 9, 0,ode, fsolve, sqrt, exp, sin, cos, arccos, pi, array, norm, solve,float,complex, dot, arange, isscalar, max, sum, flatten, shape, reshape, find, any, all, abs, plot, linspace, legend, quad, polyval,polyfit, hstack,vector, concatenate,vstack,column_stack,empty,zeros,ones,rand,vander,grid,pcolor,eig,eigs,eigvals,svd,qr,tan,det,logspace,roll,min,mean,cumsum,cumprod,diff,vectorize,lstsq,cla,eye,xlabel,ylabel,squeeze},
emphstyle=[4]\color{commandcolour},
emph={[5]__init__,__add__,__mul__,__div__,__sub__,__call__,__getitem__,__setitem__,__eq__,__ne__,__nonzero__,__rmul__,__radd__,__repr__,__str__,__get__,__truediv__,__pow__,__name__,__future__,__all__},
emphstyle=[5]\color{specmethodcolour},
emph={[6]assert,range,yield},
emphstyle=[6]\color{keywordcolour}\bfseries,
emph={[7]def, return, and, print},
emphstyle=[7]\color{black}\bfseries,
% emph={[7]self},
% emphstyle=[7]\bfseries,
literate=*%
%{:}{{\literatecolour:}}{1}%
%{=}{{\literatecolour=}}{1}%
%{-}{{\literatecolour-}}{1}%
%{+}{{\literatecolour+}}{1}%
%{*}{{\literatecolour*}}{1}%
{/}{{\literatecolour/}}{1}%
{!}{{\literatecolour!}}{1}%
%{(}{{\literatecolour(}}{1}%
%{)}{{\literatecolour)}}{1}%
%{[}{{\literatecolour[}}{1}%
%{]}{{\literatecolour]}}{1}%
{<}{{\literatecolour<}}{1}%
{>}{{\literatecolour>}}{1}%
{>>>}{{\textcolor{promptcolour}{>>>}}}{1}%
,%
breaklines=true,
breakatwhitespace= true,
xleftmargin=\framemargin,
xrightmargin=\framemargin,
aboveskip=1ex,
belowskip=1ex,
frame=trbl, %trbl
numbers=none,
%frameround=tttt,
rulecolor=\color{black!40},
%framexleftmargin=\framemargin,
%framextopmargin=.1ex,
%framexbottommargin=.1ex,
%framexrightmargin=\framemargin,
%framexleftmargin=1mm, framextopmargin=1mm, frame=shadowbox, rulesepcolor=\color{blue},#1
%frame=tb,
%backgroundcolor=\color{yellow!10}
backgroundcolor=\color{gray}
%}
}

@@

\newcommand{\inpyth}{\lstinline[style=pythonstyle, basicstyle=\ttfamily]} %[]%

\lstnewenvironment{inpython}[1][]{
\lstset{style=pythonstyle, frame=trbl, belowcaptionskip=\belowtitleskip}
}{}

@@

\title{Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations}
\author{Mikael Mortensen and Diako Darian}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
\section*{Abstract}

\section{Introduction}

\section{Preliminaries}
The Navier-Stokes equations used to describe turbulent flow in a doubly 
periodic channel can be written in rotational form as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{P}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time and $\bm{\mathcal{H}}(\bm{x}, t) = \bm{u}\times 
\bm{\omega}$ (where $\bm{\omega} = \nabla \times \bm{u}$)  represents the 
nonlinear convection.  The constant dynamic viscosity is denoted as $\nu$ and 
$P$ is a pressure modified to account for the kinetic energy, i.e., $P = p + 
\bm{u} \cdot \bm{u}/2$, where $p$ the instantaneous pressure 
normalized by a constant density. The computational domain is $\Omega=[-1, 
1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such 
that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions 
with periodic lengths $L_y$ and $L_z$, respectively.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall normal direction. Three different sets of basis functions and function spaces are relevant for the wall normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_0, \phi_1,\ldots, \phi_{N_x}\} \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_0, \D{\phi}_1, \ldots, \D{\phi}_{N_x-2} \} \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_0, \N{\phi}_1, 
\ldots, \N{\phi}_{N_x-4} \} \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
suggested by Shen, and satisfy, respectively, the boundary conditions 
$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. The 
last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in 
V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 
1) = 0 \}$.

Three-dimensional basis functions and function spaces that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, 
 \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\} 
 \label{eq:wavenumbermesh}
\end{align}
is the wavenumber mesh for $V_N$. The other two wavenumber meshes $\D{K}_N$ and 
$\N{K}_N$ are both equal to $K_N$ except from in index $l$ that 
ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
notation $K_N^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
dimensions of $K_N$.

The computational mesh in real space can be created using $N = (N_x, N_y, 
N_z)$ intervals, where the two periodic directions use uniform intervals. The 
computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

@@

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, 
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for 
periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, 
where quadrature is employed for integration. As such the scalar product 
corresponds to sums that may be computed using fast transforms. Using 
$\N{\psi}_{\bm{k}}$ for $v$ we obtain
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is used to represent the Shen scalar product in the 
$x$-direction. The weights, $w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.\footnote{Note that 
$c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq. 
(2.7) of Shen2), but for the discrete $l^2_w(\Omega)$ it follows that 
$c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}

>>>>

\documentclass[11pt, oneside]{elsarticle}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

@@

@@

@@

\begin{document}
\begin{frontmatter}

\title{Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations}
\author[mmo]{Mikael Mortensen}
\ead{mikaem@math.uio.no}
%\date{}							% Activate to display a given date or no date
\address[mmo]{Department of Mathematics, Division of Mechanics, University of Oslo}

\begin{abstract}
A fully (pseudo-)spectral solver for direct numerical simulations of large-scale turbulent channel flows is described. The solver utilizes the Chebyshev base functions suggested by J. Shen [SIAM J. Sci. Comput., 16, 1, 1995], that lead to stable and robust numerics, even at very large scale. New algorithms for direct solution of the linear systems are devised, and we show that the solver is accurate and applicable for very large simulations. Specifically, the solver is shown to reproduce first order statistics of a channel flow at $Re_{\tau}=2000$ by Hoyas and Jim\'{e}nez [Phys. Fluids, 18(1), 2006], which is higher than has been achieved in a channel by a fully spectral method before. 
\end{abstract}
\begin{keyword}
DNS \sep Fourier \sep Chebyshev
\end{keyword}

\end{frontmatter}
\section{Introduction}
Direct Numerical Simulations (DNS) of turbulent flows is a very important research tool, utilized across a range of scientific communities. DNS is used extensively to validate statistical models, and to further our understanding of complex mechanisms taking place inside turbulent flows. One of the many advantages of DNS is that it provides all information about a flow, and quantities that can be very hard to study experimentally, like velocity-pressure interactions, are trivially extracted from a DNS. In this regard, DNS both complements and extends the knowledge we are able to extract from experiments.

The most commonly known DNS use simple geometries, because turbulence physics may then most easily be isolated and studied. Isotropic and homogeneous turbulence are usually studied in triply periodic domains, which allows a spectral Fourier decomposition in all three spatial directions. The equations are solved in spectral space and nonlinearities are usually handled with pseudo-spectral techniques. An important feature is that the pressure can be trivially eliminated, which simplifies the Navier-Stokes equations substantially.

Inhomogeneous, wall bounded, flows, successfully studied with DNS are, e.g., the pipe flow, free shear layers and channel flow between two parallel plates (both Couette and Poiseuille). In this paper we will constrain ourselves to the pressure driven Poiseuille channel flow. In a channel there are two periodic directions that can be handled with Fourier expansions, whereas the wall normal direction is non-periodic and requires a different type of discretization. There are many challenges associated with this inhomogeneity not faced by the pure Fourier solvers, but the first problem at hand is the discretization. The first DNS channel solvers typically used a Chebyshev expansion for the wall-normal direction and, as such, were still able to obtain spectral accuracy in all three spatial directions. A Chebyshev-tau technique, that utilize the recurrence relations of the Chebyshev polynomials, is then used to approximate derivatives, and the coefficient matrices that appear (tridiagonal) can be inverted directly and efficiently. A downside to the Chebyshev-tau method is usually quoted \cite{canuto1988} as numerical stability and roundoff errors, caused by the recurrence relation, and severe condition numbers of the coefficient matrices. For Chebyshev-tau methods the condition numbers have been reported to grow with size as $\mathcal{O}(N^8)$, for a discretization using N points in the wall-normal direction. Discouraged by these numbers, all major recent channel flow simulations have found other, non-spectral, ways of discretizing the non-periodic direction.

The largest known channel simulations to date have been performed by Lee and Moser \cite{leemoser15}, where for $Re_{\tau}\approx 5200$ they used a computational box of resolution $[10240 \times 1530 \times 7680]$ for the streamwise, wall-normal and spanwise directions, respectively. Lee and Moser used seventh-order B-splines for the wall-normal direction. Other simulations of similar magnitude are performed by Hoyas and Jim\'{e}nez \cite{hoyas06, hoyas08}, Lozano-Duran and Jim\'{e}nez \cite{Lozano2014} and Bernardini, Pirozzoli and Orlandi \cite{bernardini2014}. Bernardini et al. used second order finite differences throughout. The Jim\'{e}nez group used dealiased Fourier in the two periodic directions and seventh order compact finite differences, with fourth-order consistency and extended spectral-like resolution \cite{Lele92}, for the wall-normal direction. The solver by Jim\'{e}nez' group is reported to switch from Chebyshev to finite differences if the resolution is above a certain threshold \cite{hoyas08} (reached around $Re_{\tau}=1000$). In other words, they attempt to use spectral methods for as large $Re_{\tau}$ as possible. Spectral methods are attractive for their accuracy, but also for their resolution properties, that are superior to those of any finite difference or spline method. As such, it is desirable to develop fully spectral solvers that can be used for large-scale turbulence simulations.

In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to make Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to fast ($\mathcal{O}(N)$) direct solvers. To the authors knowledge, the bases have not been used for large-scale channel flow simulations, and algorithms for the required direct solvers have not, until now, been devised. Shen's bases have been used for the Navier Stokes equations before, though. Bouchon et.al.  \cite{Bouchon01} describe a spectral-Galerkin formulation very similar to the one used in this paper. However, they choose the Legendre basis over the Chebyshev, due to the simpler structure of the matrices and the lack of weight functions for the Legendre basis. Also, the channel simulations performed were rather small, and there is also the issue of fast transforms for the Legendre basis.

In this paper we will describe and assess a channel flow solver based on Shen's Chebyshev basis \cite{Shen95}. We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS} and we will here describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise in Sec~\ref{sec:implementation}. We will finally show in Sec~\ref{sec:verification} that roundoff is not a major issue, and that the spectral-Galerkin method is in deed applicable to large-scale simulations, larger than what has been achieved for fully spectral channel simulations before.

\section{Preliminaries}
\label{sec:prelim}
The Navier-Stokes equations used to describe turbulent flow in a doubly 
periodic channel can be written in rotational form as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{P}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time and $\bm{\mathcal{H}}(\bm{x}, t) = \bm{u}\times 
\bm{\omega}$, where $\bm{\omega} = \nabla \times \bm{u}$, represents 
the nonlinearity.  The constant dynamic viscosity is denoted as $\nu$ 
and $P(\bm{x}, t)$ is a pressure modified to account for both the driving force, $\beta(t)$, and the kinetic energy, i.e., $P = p + \bm{u} \cdot \bm{u}/2 + \beta y$, where $p$ is the instantaneous 
pressure normalized by a constant density. The computational domain is 
$\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with channel walls 
located at $x=\pm 1$, such that no-slip applies at 
$ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions with periodic lengths $L_y$ and $L_z$, respectively.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall normal direction. Three different sets of basis functions and function spaces are relevant for the wall normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_k\}_{k=0}^{N_x} \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_k\}_{k=0}^{N_x-2} \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4} \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
suggested by Shen, and satisfy, respectively, the boundary conditions 
$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. The 
last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in 
V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 
1) = 0 \}$.

Three-dimensional basis functions and function spaces, that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, 
 \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\} 
 \label{eq:wavenumbermesh}
\end{align}
is the wavenumber mesh for $V_N$. The other two wavenumber meshes $\D{K}_N$ and 
$\N{K}_N$ are both equal to $K_N$ except from in index $l$ that 
ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
notation $K^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
dimensions of $K_N$. Likewise, ${K}^x(l)={K}_N(l, 0, 0), \D{K}^x(l)=\D{K}_N(l, 0, 0)$ and $\N{K}^x(l)=\N{K}_N(l, 0, 0)$ refer to the wavenumbers in the wall normal direction.

The computational mesh in real space can be created using $N = (N_x, N_y, 
N_z)$ intervals, where the two periodic directions use uniform intervals. The 
computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either Gauss-Lobatto or Gauss-Chebyshev points
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

@@

A three-dimensional scalar product in the weighted $L^2_w(\Omega)$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, 
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for 
periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
this work we will make use of the discrete weighted $l^2_w(\Omega)$ space, 
where quadrature is employed for integration. As such the scalar product 
corresponds to sums that may be computed using fast transforms. Using 
$\N{\psi}_{\bm{k}}$ for $v$ we obtain
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three-dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is used to represent the Shen scalar product in the 
$x$-direction. The weights, $w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.\footnote{Note that 
$c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq. 
(2.7) of \cite{Shen95}), but for the discrete $l^2_w(\Omega)$ it follows that 
$c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}

The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
complete transform. To find a transformation between physical $u$ and spectral 
$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly in the first line of (\ref{eq:sst1}) and use extensively the discrete 
orthogonality of the Fourier basis functions
\begin{align}
%\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
%\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
%\N{\psi}_{\bm{k}} \right>_w, \notag \\
\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
\sum_{(q,r,s) \in \N{K}_N} \hat{u}(q,r,s,t) \N{\psi}(q,r,s), 
\N{\psi}(l,m,n) \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t). 
           \label{eq:sst2}
\end{align}
Here $\{\N{B}_{lq}\}_{l,q=0}^{N_x-4} = (\N{\phi}_q, \N{\phi}_l)_w = 
\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
pentadiagonal mass matrix. The complete transformation is now obtained by 
setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. 
Moving to matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u} &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
 \quad \forall \, \bm{k} \in \N{K}_N, \label{eq:fstB}
\end{align}
where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that, since $\N{B}$ assembles to 
a pentadiagonal matrix, the solution ($\N{B}^{-1}$) can be obtained very fast and the complete 
transformation thus requires a fast Chebyshev transform ($\mathcal{O}(N_x \log N_x)$) and 
a fast linear algebraic solve ($\mathcal{O}(N_x)$). Similar transforms $\mathcal{T}$ 
and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
\D{\phi}_l)_w$. See the Appendix for algorithms for all required transforms.

\section{Discretization of Navier Stokes equations}
\label{sec:discretizationNS}
The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
Laplacian of the wall normal momentum equation and the curl of the momentum 
equation. Following elimination of the pressure, the equations to solve are 
\begin{align}
\frac{\partial}{\partial t} \nabla^2 u &= h_u + \nu \nabla^4 u, 
\label{eq:biharmonic} \\
\frac{\partial g}{\partial t} &= h_g + \nu \nabla^2 g, \label{eq:g} \\
f + \frac{\partial u}{\partial x} &= 0, \label{eq:f}
\end{align}
where
\begin{align}
f &= \frac{\partial v}{\partial y} + \frac{\partial w}{\partial z}, \\
g &= \frac{\partial w}{\partial y} - \frac{\partial v}{\partial z}, \\
h_u = -\frac{\partial}{\partial x} &\left( \frac{\partial 
\mathcal{H}_y}{\partial y} + \frac{\partial \mathcal{H}_z}{\partial z} \right) 
+ \left(\frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2} 
\right) \mathcal{H}_x ,
\\
h_g &= \frac{\partial \mathcal{H}_z}{\partial y} - \frac{\partial 
\mathcal{H}_y}{\partial z}.
\end{align}
The two remaining velocity components are computed from the definitions of $f$ 
and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
conditions $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
\begin{align}
	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
